\documentclass[a4paper, 11pt, twoside]{report}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% Listings package for code formatting.
\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  breaklines=true,
}


% Set up referencing.

% style=numeric - Use [1], [1,2], etc. in text
% sorting=none - Don't sort references, leave them in order referenced in text
% terseinits=true - Don't put periods after initials
% giveninits=true - Initialise given names
% minbibnames=6 - Always show at least 6 names if available
% maxbibnames=6 - Truncate lists of authors above 6 to [minbibnames] et al.
\usepackage[style=numeric, sorting=none, terseinits=true, giveninits=true, minbibnames=6, maxbibnames=6]{biblatex}

% Write surnames before given names.
\DeclareNameAlias{default}{family-given}

% Get rid of the comma after the surname.
\renewcommand*{\revsdnamepunct}{}

% Small bibliography text.
\AtBeginBibliography{\small}

% Use "References" instead of "Bibliography".
\AtBeginDocument{\renewcommand{\bibname}{References}}

% Set up reference database
\addbibresource{./bibs/introduction.bib}
\addbibresource{./bibs/background.bib}
\addbibresource{./bibs/datasets.bib}
\addbibresource{./bibs/labels.bib}
\addbibresource{./bibs/evaluation.bib}
\addbibresource{./bibs/ethics.bib}
\addbibresource{./bibs/conclusion.bib}

\usepackage{xcolor}
\usepackage{csquotes}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage[all]{nowidow}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{float}
\usepackage{subcaption}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{cleveref}
\usepackage{multirow}

\usepackage{parskip}
\usepackage{ragged2e}
\usepackage[ruled]{algorithm2e}
\usepackage{array}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{csquotes}
\usepackage{pgfgantt}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

% Used for checkmarks.
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

% Used for bullets in tables.
\newcommand{\tabitem}{~~\llap{\textbullet}~~}

\title{Semantic Synthesis of Diabetic \\ \vspace{0.1cm} Retinal Fundus Images}
\author{Joon-Ho Son}

\begin{document}
\input{title/title.tex}

% Make chapter headings smallers to save space.
\makeatletter
\def\@makechapterhead#1{%
  {\parindent \z@ \raggedright \normalfont
    \ifnum \c@secnumdepth >\m@ne
        \huge\bfseries \@chapapp\space \thechapter
        \par\nobreak
        \vskip 15\p@
    \fi
    \interlinepenalty\@M
    \Huge \bfseries #1\par\nobreak
    \vskip 30\p@
  }}
\def\@makeschapterhead#1{%
  {\parindent \z@ \raggedright
    \normalfont
    \interlinepenalty\@M
    \Huge \bfseries  #1\par\nobreak
    \vskip 30\p@
  }}
\makeatother

\begin{abstract}
Retinal fundus imaging is a modality that’s quick and easy to obtain, however its usefulness is limited by the availability of human graders.
With advancements in deep learning, data-driven models may facilitate the process of diagnosis, alleviating this bottleneck. 
Despite this, training robust machine learning models requires manually annotated labels en-masse, which are costly to obtain, and remains to be one of the primary obstacles in producing generalisable models. 
This challenge can be viewed in two aspects: the scarcity of manually annotated ground truths such as lesion segmentation maps, and class imbalance in available datasets.
Both of these can be seen in datasets designed for classification and segmentation tasks related to diabetic retinopathy (DR).

In this project, we propose a novel two-step process for generating photo-realistic fundus images conditioned on synthetic ``ground truth'' semantic labels.
First, synthetic semantic labels for retina structures and lesions are generated from random noise, conditioned on DR severity. 
These are then subsequently used to condition an image-to-image translation model to create the final retinal fundus image.
We compare the performance of these sophisticated generative models against a simple data augmentation method.

We go on to demonstrate its potential to improve performance on downstream tasks in classifying and segmenting diabetic retinal fundus images. 
\end{abstract}

\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
Firstly, I would like to express my gratitude to my supervisor \textbf{Dr Benjamin Hou} for his unwavering support and guidance throughout this project.
I would also like to thank my second marker \textbf{Dr Amir Alansary} for his valuable feedback during the early stages.

I reserve special thanks for my mother, whose countless sacrifices made a better life possible for her sons.
\end{abstract}

\tableofcontents
% \listoffigures
% \listoftables

\input{introduction/introduction.tex}
\input{background/background.tex}
\input{datasets/datasets.tex}
\input{labels/labels.tex}
\input{retinas/retinas.tex}
\input{evaluation/evaluation.tex}
\input{conclusion/conclusion.tex}
\input{appendix/appendix.tex}

\printbibliography

\end{document}